---
title: "Mapping Origin-Destination Employment in Tacoma, WA (2019)"
author: "Andrew Baker, Eman Robi, James Lyou, Alberto Melendez, Jamie Marken"
date: "January 28, 2019"
output: html_document
---

### Report
We chose Tacoma as our area of interest because of Tacoma's large shipping industry. Our data comes from the city of Tacoma GIS portal for the pierce county shapefile and the U.S. Census Bureau website for the washiington work/home tracts. We had a few issues regarding the formart of our data, for example; merging our data frames was difficult because sometimes the geo id needed to be put into certain data frames.

Our main finding is that people making less money tend to be centered closer to the city of Tacoma then further away. Also that people tend to have work in many different area. not just clsoe to where they live.

```{r Packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(dplyr)
library(tmap)
library(leaflet)
library(ggplot2)
library(shiny)
```

```{r, include=TRUE}
baseData=read.csv("./lab 2 data/OD/wa_od_main_JT01_2015.csv")
odData=baseData
#Filter out Tacoma Codes and pretty up the geocodes
odData <- filter(odData, w_geocode > 530530603000000 & w_geocode < 530530620999999 | h_geocode > 530530603000000 & h_geocode < 530530620999999)
odData$w_geocode<-as.character(odData$w_geocode)
odData$h_geocode<-as.character(odData$h_geocode)
odData$w_geocode[which(length(odData$w_geocode) == 14)]<-paste0("0",odData$w_geocode[which(length(odData$w_geocode) == 14)])
odData$h_geocode[which(length(odData$h_geocode) == 14)]<-paste0("0",odData$h_geocode[which(length(odData$h_geocode) == 14)])
#Making Columns for worktract and hometract
odData$wt<-substr(odData$w_geocode,1,11)
odData$ht<-substr(odData$h_geocode,1,11)
write.csv(odData, "od2015_Tacoma.csv")
```

```{r, include=TRUE}
racBaseData=read.csv("./lab 2 data/RAC/wa_rac_S000_JT01_2015.csv")
racData=racBaseData
racData$h_geocode[which(length(racData$h_geocode) == 14)]<-paste0("0", racData$h_geocode[which(length(racData$h_geocode) == 14)])
racData$ht<-substr(racData$h_geocode,1,11)
 
aggRACData <- select(racData, ht)
aggRACData <- unique(aggRACData)
aggRACData <- arrange(aggRACData, ht)
counter = 2
while(counter <= 42){
 aggRACData[c(colnames(racData)[counter])] <- 0
 counter <- counter + 1
}
aggRACCount <- 1
# for(i in aggRACData$ht) {
#  baseRACCount <- 1
#  while(baseRACCount <= nrow(racData)) {
#    if(racData[baseRACCount, 44] == i){
#      counterJ = 2
#      while(counterJ <= 42) {
#        if(racData[baseRACCount, counterJ] != 0) {
#          aggRACData[aggRACCount, counterJ] <- aggRACData[aggRACCount, counterJ] + racData[baseRACCount, counterJ]
#        }
#        counterJ <- counterJ + 1
#      }
#    }
#    baseRACCount <- baseRACCount + 1
#  }
#  aggRACCount <- aggRACCount + 1
# }
# write.csv(aggRACData, "rac2015_Tacoma.csv")
```

```{r, include=TRUE}
wacBaseData=read.csv("./Lab 2 Data/WAC/wa_wac_S000_JT01_2015.csv")
 
wacData=wacBaseData
wacData$w_geocode[which(length(wacData$w_geocode) == 14)]<-paste0("0", wacData$w_geocode[which(length(wacData$w_geocode) == 4)])
wacData$wt<-substr(wacData$w_geocode,1,11)
 
aggWACData <- select(wacData, wt)
aggWACData <- unique(aggWACData)
aggWACData <- arrange(aggWACData, wt)
 
counter = 2
while(counter <= 52){
 aggWACData[c(colnames(wacData)[counter])] <- 0
 counter <- counter + 1
}
aggWACCount <- 1
# for(i in aggWACData$wt) {
#   baseWACCount <- 1
#   while(baseWACCount <= nrow(wacData)) {
#     if(wacData[baseWACCount, 54] == i){
#      counterJ = 2
#      while(counterJ <= 52) {
#        if(wacData[baseWACCount, counterJ] != 0) {
#          aggWACData[aggWACCount, counterJ] <- aggWACData[aggWACCount, counterJ] + wacData[baseWACCount, counterJ]
#        }
#        counterJ <- counterJ + 1
#      }
#     }
#     baseWACCount <- baseWACCount + 1
#  }
#   aggWACCount <- aggWACCount + 1
# }
# write.csv(aggWACData, "wac2015_Tacoma.csv")
```

#Data 
```{r, include = TRUE}
ODDATA <- read.csv("od2015_Tacoma.csv")
block_data <- read.csv("tacoma_blocks2.csv")
track_data <- read.csv("tacoma_tracts2.csv")
WACT <- read.csv("wac2015_Tacoma.csv")
WACB <- read.csv("wac2015_Tacoma_Block.csv")
tacomablockshp <- st_read("tacomablocks-polygon.shp")
tacomatrackshp <- st_read("pie10ct.shp")
Washingtontrackshp <- st_read("cb_2017_53_tract_500k.shp")
```

#Part 3.2 
```{r, include = TRUE}
library(tidyverse)
baseData=read.csv("./Lab 2 Data/OD/wa_od_main_JT01_2015.csv")
odData=baseData
#Filter out Tacoma Codes and pretty up the geocodes
odData <- filter(odData, w_geocode > 530530603000000 & w_geocode < 530530620999999 | h_geocode > 530530603000000 & h_geocode < 530530620999999)
odData$w_geocode<-as.character(odData$w_geocode)
odData$h_geocode<-as.character(odData$h_geocode)
odData$w_geocode[which(length(odData$w_geocode) == 14)]<-paste0("0",odData$w_geocode[which(length(odData$w_geocode) == 14)])
odData$h_geocode[which(length(odData$h_geocode) == 14)]<-paste0("0",odData$h_geocode[which(length(odData$h_geocode) == 14)])
#Making Columns for worktract and hometract
odData$wt<-substr(odData$w_geocode,1,11)
odData$ht<-substr(odData$h_geocode,1,11)
#odData
#Plots compare Total Numbers of Jobs to Number of jobs in Goods Producing industry sectors and Number of jobs in Trade, Transportation, and Utilities industry sectors
p <-ggplot(data = odData, aes(odData$S000, odData$SI01))
p + geom_bar(stat="identity", aes(fill = odData$SI02)) + xlab("Total Number of Jobs") + ylab("Number of jobs in Goods Producing industry sectors") + labs(fill = "Number of jobs in Trade, Transportation, and Utilities industry sectors")
```


#Part 3.3
```{r, include = TRUE}
library(dplyr)

###Create Map 1
# Block Shapefile with WAC 
WACB$GEOID10= WACB$w_geocode


WACBLOCK <- merge(tacomablockshp, WACB, by=c("GEOID10"))

#Change label name for plot Block 
WACBLOCK$Population = WACBLOCK$CE01


#plot block
BLOCK_MAP <- tm_shape(WACBLOCK) +tm_polygons() +tm_symbols(col="Population") + tm_borders() +tm_layout(title = "Number of jobs with earnings $1250/month or less in the City of Tacoma on a Block Level", title.position = c("RIGHT", "TOP")) 

BLOCK_MAP

##Create Map 2
#Number of jobs with earnings $1250/month or less 
#plot 2 track
WACT$GEOID10=WACT$wt

WACTRACK <- merge(tacomatrackshp, WACT, by=c("GEOID10"))
WACTRACK$Population = WACTRACK$CE01

TRACK_MAP <- tm_shape(WACTRACK) +tm_polygons() +tm_symbols(col="Population") + tm_borders() +tm_layout(title = "Number of jobs with earnings $1250/month or less in the City of Tacoma on a Tract Level", title.position = c("RIGHT", "TOP"))

TRACK_MAP
```

#Part 3.4
```{r, include = TRUE}
###Create Map of OD 
HomeTrack1 <-filter(ODDATA, ht == 53053060905)

Washingtontrackshp$wt = Washingtontrackshp$GEOID

HomeTrack1Merge <- merge(Washingtontrackshp, HomeTrack1, by=c("wt"), all = TRUE)
HomeTrack1Merge[is.na(HomeTrack1Merge)] <- 0 

 tm_shape(HomeTrack1Merge) +tm_polygons() +  tm_symbols(col="wt") + tm_borders() +tm_layout(title = "Points of places where people who live in area 53053060905 work", title.position = c("RIGHT", "TOP")) 

```

### for Phillip Neel
We know that our assignment is incomplete, we worked very hard to finish this assignment, but we faced alot of difficulty around mapping our data spatially. Especially at part 3.4. We will try to do better next time. 
